---
title: "Assignment 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r values, echo = FALSE, include= FALSE}
source("../source/analysis.R", print.eval = TRUE)
library(readr)
library("tidyverse")
library("tidyr")
library(ggplot2)
library("plyr")
library(usdata)
library("dplyr")
```

## Introduction

In this project, we are analyzing a wide set of incarceration data at the county and state levels. Specifically, the data includes various races and their representation in various facets of jail. I chose to analyze male, black and white, and county specific data in calculating my values.

The first value I calculated was the average county-wide percentage of males in prison in 2016 (the most recent year of reliable data). I found that `r value_1*100` percent of individuals in the average county are incarcerated as of 2016. I found this interesting as it could be useful to assess what a standard jailed percentage would be in some later values (values 4 and 5).

The second value I calculated was the percentage of the total population that was white in 2016 (specifically adults aged 15 to 64). The percentage winded up being `r value_2*100`. I found this interesting because I was curious how much of the United States was white, and compare that to my anecdotal experience in Seattle. 

The third value I calculated was the white percentage of the total jail population in 2016, which was `r value_3*100`. I found this interesting as it did not match the percentage of white population. This data meant white people are not proportionally represented in jail. 

The fourth value I calculated was the percent of King County that was in jail in 2016, which ended up being `r value_4*100`. I found this interesting as it was slightly different than the `r value_1*100` we found to be the standard for an average county in Value 1. 

Finally, I did the same calculation as Value 4, but for my home county of Santa Clara County. The percentage incarceration in Santa Clara County was `r value_5*100` in 2016. Again, this differs from the national average (and from King County for that matter).

### Trends Over Time Chart

```{r chart1, echo = FALSE, message = FALSE}
source("../source/Chart1.R", print.eval = TRUE)
```

This chart shows how the black and white population each are represented in jail separately. The clear takeaway is the significant difference in the two lines. The white population has historically and continually had much lower representation in jail relative to the national population. Another minor pattern is the increase in black incarceration and eventually decrease in the mid 2000s. However, the significant takeaway still is the divide that has continued to remain stable throughout the last 30  years. 


### Variable Comparison Chart

```{r chart2, echo = FALSE, message = FALSE}
source("../source/Chart2.R", print.eval = TRUE)
```

This chart was included since it signifies the difference in racial demographics in large states. Clearly, the white population heavily outweighs the black population is nearly all of the 13 largest states (choosing 13 since Washington is the 13th largest state). It goes to show that the US is still majority white, even if incarceration rates don't follow that trend... 


### Map Chart

```{r chart3, echo = FALSE, message = FALSE}
source("../source/Chart3.R", print.eval = TRUE)
```

This map shows how different states have faired in narrowing the divide demonstrated in the first chart. Clearly, Texas and Missouri (among a few other states) have some serious problems when it comes to narrowing the divide between white and black representation in prison, with both at or around and 8 basis point different in the respective percentage of blacks and whites in prison. 
